Faithful Variable Screening for\\ High-Dimensional Convex Regression

### 2014-10-25 tex/scam/summary.tex
tried to fix the abrupt changes

The starting point for our approach is the observation that least squares
nonparametric estimation under convexity constraints is equivalent to
a finite dimensional quadratic program.  Specifically, the infinite
dimensional optimization 
\begin{align}
\begin{split}
\text{minimize} & \quad \sum_{i=1}^n (Y_i - f(\x_i))^2 \\
\text{subject to} &  \quad f:\reals^p\rightarrow\reals\ \text{is
  convex}
\end{split}
\end{align}
is equivalent to the finite dimensional quadratic
program 
\begin{align}
\label{eq:convreg}
\begin{split}
\text{minimize}_{f, \beta} & \;\; \sum_{i=1}^n (Y_i - f_i)^2 \\
\text{subject to} & \;\; f_j \geq f_i + \beta_i^T (\x_j-\x_i),\; \text{for
    all $i,j$}.
\end{split}
\end{align}
See \cite{Boyd04}, Section 6.5.5.
Here $f_i$ is the estimated function value $f(\x_i)$, and the vectors
$\beta_i \in \reals^d$ represent supporting hyperplanes to the
epigraph of $f$.  Importantly, this finite dimensional quadratic program does
not have tuning parameters for smoothing the function.

For general regression, using an additive approximation for variable
selection may make errors.  In particular, the nonlinearities in the
regression function may result in an additive component being wrongly
zeroed out.  We show that this will not happen for convex regression
under appropriate conditions.

>>>>

The starting point for our approach is the observation that least squares
nonparametric estimation under convexity constraints is equivalent to
a finite dimensional quadratic program.  Specifically, the infinite
dimensional optimization 
\begin{align}
\begin{split}
\text{minimize} & \quad \sum_{i=1}^n (Y_i - f(\x_i))^2 \\
\text{subject to} &  \quad f:\reals^p\rightarrow\reals\ \text{is
  convex}
\end{split}
\end{align}
is equivalent to the finite dimensional quadratic
program 
\begin{align}
\label{eq:convreg}
\begin{split}
\text{minimize}_{f, \beta} & \;\; \sum_{i=1}^n (Y_i - f_i)^2 \\
\text{subject to} & \;\; f_j \geq f_i + \beta_i^T (\x_j-\x_i),\; \text{for
    all $i,j$}.
\end{split}
\end{align}
Here $f_i$ is the estimated function value $f(\x_i)$, and the vectors
$\beta_i \in \reals^d$ represent supporting hyperplanes to the
epigraph of $f$.  See \cite{Boyd04}, Section 6.5.5.
Importantly, this finite dimensional quadratic program does
not have tuning parameters for smoothing the function.

This formulation of convex regression is subject to the curse
of dimensionality.  Moreover, attempting to select variables
by regularizing the subgradient vectors $\beta_i$ with
a group sparsity penality is not effective.  Intuitively, the reason
is that all $p$ components of the subgradient $\beta_i$
appear in every convexity constraint 
$ f_j \geq f_i + \beta_i^T (\x_j-\x_i)$; small changes to the 
subgradients may not violate the constraints.  Experimentally,
we find that regularization with a group sparsity penality
will make the subgradients of irrelevant variables 
small, but may not zero them out completely.

This motivates us to consider an additive approximation. 
Under a convex additive model, each component of the subgradient
appears only in the convexity constraint for
the corresponding variable:
\begin{equation}
f_{ki'} \geq f_{ki} + \beta_{ki}(x_{ki'}-x_{ki}) 
\end{equation}
where $f_{ki} = f_{k}(x_{ki})$ and $\beta_{ki}$ is the
subgradient at point $x_{ki}$.   As we show, this leads to
an effective variable selection procedure.  The
shape constraints play an essential role.
For general regression, using an additive approximation for variable
selection may make errors.  In particular, the nonlinearities in the
regression function may result in an additive component being wrongly
zeroed out.  We show that this cannot happen for convex regression
under appropriate conditions.
